---
title: "gtsummary exercise"
author: "Alex Shen"
date: "2022/7/3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message = F, error = F, warning = F)
```
Normally a data scientist will go through the following steps for their job:

* import data
* clean data
* explore data
* analyze data
* reporting data

## Loading data

```{r loading package/data}
library(tidyverse)
library(gtsummary)
data(trial)
```

## Looking at data
 
```{r looking data}
trial
```



```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 
```


Age density comparison between two groups
```{r}
trial %>% drop_na(age) %>% 
ggplot(aes(x=age, col = trt)) + geom_density()+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
```

Mean and SD for each group
```{r}
trial %>%  
  group_by(trt) %>% 
  summarize(avg.age = mean(age, na.rm = T), sd=sd(age,na.rm=T),count = n())
```


Relationship between 2 categorical variables
```{r}
trial %>%
  drop_na(response) %>% 
  ggplot(aes(x = trt, fill = factor(response))) +
  geom_bar(position = "fill",alpha = 0.5)
```


Relation of of one numerical and two categorical variable
```{r}
ggplot(trial, aes(x = trt, y = age, fill = stage)) +
  geom_boxplot()
```

Journal level presentation of data in table

table 1 with 1 group

```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary1 <- sm_trial %>% 
  select(-trt) %>% 
  tbl_summary()

tbl_summary1
  
```
"Table 1" with 2 groups

```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary2 <- sm_trial %>% 
  tbl_summary(
    by = trt,
    statistic = response ~ "{n}/{N}({p}%)",
    label = grade ~ "Pathologic tumor grade",
    digits = age ~ 2
  )

tbl_summary2
  
```


```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary3 <- sm_trial %>% 
  tbl_summary(
    by = trt,
    statistic = list(all_continuous() ~ "{mean}({sd})",
                     response ~ "{n}/{N}({p}%)"),
    label = list(grade ~ "Pathologic tumor grade", age ~ "Patient age(year)"),
    digits = age ~ 2
  )

tbl_summary3
  
```


```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary4 <- sm_trial %>% 
  tbl_summary(
    by = trt,
    statistic = list(all_continuous() ~ "{mean}({sd})",
                     response ~ "{n}/{N}({p}%)"),
    label = list(grade ~ "Pathologic tumor grade", age ~ "Patient age(year)"),
    digits = age ~ 2
  ) %>% 
  add_p()
  

tbl_summary4
  
```


```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary5 <- sm_trial %>% 
  tbl_summary(
    by = trt,
    missing = "no",
  ) %>% 
  add_p() %>% 
  add_overall() %>% 
  add_n() %>% 
  add_stat_label(
    label = all_categorical() ~ "n/N(%)"
  )
  

tbl_summary5
  
```




```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary6 <- sm_trial %>% 
  tbl_summary(
    by = trt,
    missing = "no",
  ) %>% 
  add_p() %>% 
  add_overall() %>% 
  add_n() %>% 
  add_stat_label(
    label = all_categorical() ~ "n/N(%)"
  ) %>% 
  bold_labels() %>% 
  italicize_levels() %>% 
  bold_p(t = 0.8)
  

tbl_summary6
  
```

```{r}

iris %>% 
    group_by(Species) %>% 
    mutate(Count = n()) %>%
    group_by(Species, Count) %>%
    summarize_at(vars(), funs(mean, sd))
```


