---
title: "Class one for biometrics"
author: "Alex Shen"
date: "2022/7/13"
output: html_document
---

---
title: "An Introduction to R for Jixing Biometrics"
author: "Alex Shen"
date: '2022-07-13'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R learning vision

All for the purpose of use in our work, i.e., in **data management**, **medical monitoring**, and **statistical programming**.

-   We can use it for our actual work;
-   Or use it as a tool to monitor the work we outsourced to CRO

### A bar chart use case

Here is the youtube video worth looking<https://www.youtube.com/watch?v=TN10L3QrBXQ&t=940s>


*Create mean Sepal.length for each species*
```{r}
library(dplyr)
data <- iris %>% 
  group_by(Species) %>% 
  summarize(
    mean_Sepal.Length = mean(Sepal.Length),
    .groups = "drop"
  )
```


*Static visualization by ggplot*
```{r}
library(ggplot2)
data |>
  ggplot(
    aes(x = Species,
        y = mean_Sepal.Length)
  ) +
  geom_bar(stat = "identity")
```

Interactive visualization via plotly

```{r}
library(plotly) 
data %>% 
  plot_ly(
    x=~Species,
    y=~mean_Sepal.Length,
    type = "bar"
  )
```

Interactive visualization via echarts4r. A bar chart with tooltip trigger of "item".
```{r}
library(echarts4r)

data %>% 
  e_charts( x = Species ) %>% 
  e_bar(serie = mean_Sepal.Length) %>% 
  e_tooltip(trigger = "item")
```

### A line chart use case in echarts4r

```{r}
dates<-seq.Date(
  from = as.Date("2021-01-01"),
  by = "day",
  length.out = nrow(iris)/3
  )
  
data <-iris %>% 
  mutate(
    Date = rep(dates,3)
    ) %>% 
  select(Date, Petal.Width, Species)
```

```{r}
data %>% 
  group_by(Species) %>% 
  e_charts(x = Date) %>% 
  e_line(serie = Petal.Width) %>% 
  e_tooltip(trigger = "axis")
```


### Statistical programming use case (gtsummary)

#### Step 1 Loading simulated trial data

```{r loading package/data}
library(tidyverse)
library(gtsummary)
data(trial)
```

#### Step 2 Select 4 variables to form a small dataset

```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 
```

Age density comparison between two groups
```{r}
trial %>% drop_na(age) %>% 
ggplot(aes(x=age, col = trt)) + geom_density()+
  theme_bw()+
  theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())
```

Mean and SD for each group
```{r}
trial %>%  
  group_by(trt) %>% 
  summarize(avg.age = mean(age, na.rm = T), sd=sd(age,na.rm=T),count = n())
```

Relationship between 2 categorical variables
```{r}
trial %>%
  drop_na(response) %>% 
  ggplot(aes(x = trt, fill = factor(response))) +
  geom_bar(position = "fill",alpha = 0.5)
```


Relation of of one numerical and two categorical variable
```{r}
ggplot(trial, aes(x = trt, y = age, fill = stage)) +
  geom_boxplot()
```

Journal level presentation of data in table

table 1 with 1 group


```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary1 <- sm_trial %>% 
  select(-trt) %>% 
  tbl_summary()

tbl_summary1
  
```
"Table 1" with 2 groups

```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary2 <- sm_trial %>% 
  tbl_summary(
    by = trt,
    statistic = response ~ "{n}/{N}({p}%)",
    label = grade ~ "Pathologic tumor grade",
    digits = age ~ 2
  )

tbl_summary2
  
```


```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary3 <- sm_trial %>% 
  tbl_summary(
    by = trt,
    statistic = list(all_continuous() ~ "{mean}({sd})",
                     response ~ "{n}/{N}({p}%)"),
    label = list(grade ~ "Pathologic tumor grade", age ~ "Patient age(year)"),
    digits = age ~ 2
  )

tbl_summary3
  
```


```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary4 <- sm_trial %>% 
  tbl_summary(
    by = trt,
    statistic = list(all_continuous() ~ "{mean}({sd})",
                     response ~ "{n}/{N}({p}%)"),
    label = list(grade ~ "Pathologic tumor grade", age ~ "Patient age(year)"),
    digits = age ~ 2
  ) %>% 
  add_p()
  

tbl_summary4
  
```



```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary4 <- sm_trial %>% 
  tbl_summary(
    by = trt,
    statistic = list(all_continuous() ~ "{mean}({sd})",
                     response ~ "{n}/{N}({p}%)"),
    label = list(grade ~ "Pathologic tumor grade", age ~ "Patient age(year)"),
    digits = age ~ 2
  ) %>% 
  add_p()
  

tbl_summary4
  
```


```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary5 <- sm_trial %>% 
  tbl_summary(
    by = trt,
    missing = "no",
  ) %>% 
  add_p() %>% 
  add_overall() %>% 
  add_n() %>% 
  add_stat_label(
    label = all_categorical() ~ "n/N(%)"
  )
  

tbl_summary5
  
```




```{r}
sm_trial <- trial %>% 
  select(trt, age, grade, response) 

tbl_summary6 <- sm_trial %>% 
  tbl_summary(
    by = trt,
    missing = "no",
  ) %>% 
  add_p() %>% 
  add_overall() %>% 
  add_n() %>% 
  add_stat_label(
    label = all_categorical() ~ "n/N(%)"
  ) %>% 
  bold_labels() %>% 
  italicize_levels() %>% 
  bold_p(t = 0.8)
  

tbl_summary6
  
```



```{r}

iris %>% 
    group_by(Species) %>% 
    mutate(Count = n()) %>%
    group_by(Species, Count) %>%
    summarize_at(vars(), funs(mean, sd))
```









 
```{r looking data}
trial
```

## Learning machanism

-   Recommend to watch video first, take shared learning the second, and do the assignment and practice the next.
-   Teacher and Learners switch approach to speed up learning, fine-tuned to what matters, and promote dicussion and brainstorming
-   Goal is to use it in 2 months? for Nadia to use it for data cleaning first, and Shenglin to compare table generation in SAS and R.
-   Gathered together every week? or 2 weeks?

## Some of topics relevant to us(Our sylabus)

-   Understand and use R studio
-   Import data into R: read.csv, read_csv from readr, read_sas from haven
-   Perform data manipulation using dpyr and tidyr packages
-   Static and interactive visulization using ggplot2, plotly, and echart4r
-   Table generation using gt, dt, gtsummary et al.
-   R shiny dashboard using flexdashboard and shinydashboard

## Some of the useful resource for R learning

-   In R studio help
-   Online books
    -   [Data Visualization](https://rkabacoff.github.io/datavis/)
    -   [R blogger](https://www.r-bloggers.com/)
    -   [R for Data Science](https://r4ds.had.co.nz/)
